---
output: github_document
---
<!-- badges: start -->
[![R-CMD-check](https://github.com/pedroguarderas/CFINI/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/pedroguarderas/CFINI/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->
  
<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```
CFINI
=====

## Introduction
This package was created with implement different computational routines that can be applied to 
solve pricing problems, that are usually presented in Finance and Actuarial Science.

Here a list of the main numerical methods implemented inside CFINI. Many of this functions are 
coded with C++ by employing the packages Rcpp and RcppEigen.

- Pricing with trees
- Pricing with multinomial trees
- Ordinary differential equation solver implemented with the predictor-corrector method
- Diffusion solver with Euler implicit scheme
- Diffusion solver with Crank-Nicolson scheme

The different methods in the CFINI package are implemented based on the concepts and algorithms
introduced in the following sources of information:

- The two volume book [Stochastic Calculus for Finance I and II](https://doi.org/10.1007/978-0-387-22527-2) 
of Steven E. Shreve.
- The book [Finite difference methods in financial engineering](
https://www.wiley.com/en-us/Finite+Difference+Methods+in+Financial+Engineering:+A+Partial+Differential+Equati
on+Approach-p-9781118856482) of Daniel J. Duffy.
- The specialization of coursera [Financial Engineering and Risk Management](https://www.coursera.org/specializations/financialengineering) a MOOC provided by Columbia University.
- The MOOC of coursera [Interest Rate Models](https://www.coursera.org/learn/interest-rate-models) 
provided by EPFL.

## Installation
To install the package **CFINI** directly from github, you can proceed in the following way making 
use of the devtools library
```{r, eval=FALSE}
library( devtools )
install_github( "pedroguarderas/CFINI" )
```
