---
title: "Black Scholes model"
author: "Pedro Guarderas"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Black Scholes model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

<!-- move to a theory appendix? -->

A general representation of the equation related to the Black-Scholes model:

\begin{equation}
\left\{
\begin{array}{ll}
\dfrac{\partial V}{\partial t}( t, S ) +
\alpha( t, S ) \dfrac{\partial^2 V}{\partial S^2}( t, S ) +
\beta( t, S ) \dfrac{\partial V}{\partial S}( t, S ) + 
\gamma( t, S ) V( t, S ) = 0 & \forall ( t, S ) \in [0,T] \times [ S_l, S_h ] \\
\text{FC:} & \\
V( T, S ) = V_0( S ) & \forall S \in [ S_l, S_h ] \\
\text{BC:} & \\
V( t, S_l ) = v_1( t ) & \forall t \in [ 0, T ] \\
V( t, S_h ) = v_2( t ) & \forall t \in [ 0, T ]
\end{array} 
\right.
\end{equation}

<!-- ?? the particular ...? -->
the particular case of Black-Scholes is given by 
$\alpha( t, S ) = \frac{1}{2}\sigma^2 S^2$, $\beta( t, S ) = r S$ and $\gamma( t, S ) = -r$
where $\sigma, r$ are constants. 

Note that this model considers a final condition (*FC*) which usually represents the value of the
option at the end of the contract. 

<!-- second? -->
<!-- relates? -->
<!-- See pages 83 and 88 of  -->
<!-- Capiński, Zastawniak, *Mathematics for ﬁnance : an introduction to ﬁnancial engineering* (ISBN 1-85233-330-8) -->
The idea of being able to produce a differential model that will be capable of producing a
consistent pricing model to hedge the final option is supported by the the second fundamental
theorem of asset pricing. This theorem precisely relates the ability to hedge arbitrary claims to
the uniqueness of martingale measures.

**Transformation of the particular case of Black-Scholes**

<!-- ?? right change -->
The previous model can be transformed to the *classical diffusion* problem by the right change of
variables:

\begin{equation}
V( t, S ) = e^{\alpha x + \beta \tau} u( x, t )\\
\alpha = -\frac{1}{2}\left( \frac{2r}{\sigma^2}  - 1 \right) \\
\beta = -\frac{1}{4}\left( \frac{2r}{\sigma^2} + 1 \right)^2 \\
S = e^x \\
t = T - \frac{2 \tau}{\sigma^2}
\end{equation}


## Example

1. Load required packages

```{r,eval=TRUE,message=FALSE, warning=FALSE}
library( CFINI )
library( plotly )
```

2. Related coefficients
```{r,eval=TRUE,message=FALSE, warning=FALSE}
# Diffusion parameter constant
Nt <- 300
Nx <- 150
alpha <- matrix( 10^(-2.3), Nt, Nx )
theta <- 0.25
```

3. Time grid
```{r,eval=TRUE,message=FALSE, warning=FALSE}
t0 <- 0
t1 <- 1
t <- cf_uniform_grid( t0, t1, Nt )
```

4. Value grid
```{r,eval=TRUE,message=FALSE, warning=FALSE}
x0 <- -0.5
x1 <- 0.5
x <- cf_uniform_grid( x0, x1, Nx )
```

5. Initial conditions
```{r,eval=TRUE,message=FALSE, warning=FALSE}
If <- function( x ) if ( abs(x) <= 0.1 ) return( 1 ) else return( 0 )
If <- Vectorize( If )
I <- sapply( x, FUN = If )
```

6. Boundary conditions
```{r,eval=TRUE,message=FALSE, warning=FALSE}
A <- rep( 0, Nt )
B <- rep( 0, Nt )
```

<!-- ?? This is theory? move to another section? restart numbering? -->

\begin{equation}
A = \begin{bmatrix}
b_1 & a_1 & 0 & 0 & \cdots & 0 \\
c_1 & b_2 & a_2 & 0 & \cdots & 0 \\
0 & c_2 & b_3 & a_3 & \cdots & 0 \\
\vdots & \vdots & \ddots & \ddots & \ddots & \vdots \\
0 & 0 & \cdots & c_{n-2} & b_{n-1} & a_{n-1} \\
0 & 0 & \cdots & 0 & c_{n-1} & b_n
\end{bmatrix}
\end{equation}

<!-- ?? provides an equation of form -->
5. The implicit Euler method has the following form for the equation
\begin{equation}
u_{n+1,i} - u_{n,i} = \lambda_{n, i} ( u_{n+1,i+1} - 2 u_{n+1,i} + u_{n+1,i-1} ) +
\rho_{n,i} ( u_{n+1,i+1/2} - u_{n+1,i-1/2} ) + 
\gamma_{n,i} u_{n+1,i}
\end{equation}

<!-- ?? -->
<!-- From this equation, -->
From the previous scheme, at every time step $n$ we formulate a tridiagonal problem $A u_n = d$,
with the folowing definitions:

\begin{eqnarray}
\lambda_{n, i} & = & \alpha_{n, i} \frac{\Delta t_n}{\Delta x_{i}\ \Delta x_{i+1}} \\
\rho_{n, i} & = & \beta_{n, i} \frac{\Delta t_n}{2(x_{i+1} - x_{i})} \\
a_i & = & -\lambda_{n,i} \\
b_i & = & 1 + 2 \lambda_{n,i} \\
c_i & = & -\lambda_{n,i} \\
d_i & = & u_{n,i}
\end{eqnarray}

```{r,eval=TRUE,message=FALSE, warning=FALSE}
Ueu <- cf_diff_solv_euls( alpha, I, A, B, t, x, FALSE )
```

6. Solving with Crank-Nicolson method

\begin{equation}
u_{n+1,i} + \theta (  \lambda^2_{n,i} \Delta_x u_{n+1,i+1} -  \lambda^1_{n,i} \Delta_x u_{n+1,i} ) =
u_{n,i} + ( 1 - \theta ) (  \lambda^2_{n,i} \Delta_x u_{n,i+1} -  \lambda^1_{n,i} \Delta_x u_{n,i} )
\end{equation}

\begin{eqnarray}
\lambda^1_{n,i} & = & \alpha_{n, i} \frac{\Delta t_n}{\Delta x_{i}\ \Delta x_{i+1}} \\
\lambda^2_{n,i} & = & \alpha_{n, i} \frac{\Delta t_n}{\Delta x_{i+1}\ \Delta x_{i+1}} \\
a_i & = & -\theta \lambda^1_{n,i} \\
b_i & = & 1 + \theta ( \lambda^1_{n,i} + \lambda^2_{n,i} ) \\
c_i & = & -\theta \lambda^2_{n,i} \\
d_i & = & u_{n,i} + ( 1 - \theta ) (  \lambda^2_{n,i} \Delta_x u_{n,i+1} -  \lambda^1_{n,i} \Delta_x u_{n,i} )
\end{eqnarray}

```{r,eval=TRUE,message=FALSE, warning=FALSE}
Ucn <- cf_diff_solv_cns( theta, alpha, I, A, B, t, x, FALSE )
```

```{r,eval=TRUE,message=FALSE, warning=FALSE}
alph <- max( alpha )
phi <- function( x ) dnorm( x, mean = 0, sd = sqrt( 2 * alph * t[ Nt ] ) )
s <- sapply( x, FUN = function( y ) integrate( function( x, y ) If( y - x ) * phi( x ), -Inf, Inf, y )$value )

eeu <- matrix( Ueu$u[Nt,] - s, Nx, 1 ) # Euler error
ecn <- matrix( Ucn$u[Nt,] - s, Nx, 1 )

```

7. Plotting results
```{r,eval=TRUE,fig.show='hold', fig.show='hold', fig.align="center", fig.width=7, fig.height=4}
plot_ly( x = Ucn$t, y = Ucn$x, z = Ueu$u, alpha = 0.8 ) %>% 
  layout( scene = list( xaxis = list(title = "t"),
                        yaxis = list(title = "S"),
                        zaxis = list(title = "u") ) ) %>%
  add_surface()
```
