---
title: "Diffusion solver"
author: "Pedro Guarderas"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Diffusion Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Introduction

1. Loading packages
```{r,eval=TRUE,message=FALSE, warning=FALSE}
library( CFINI )
library( plotly )
```
\begin{equation}
\left\{
\begin{array}{ll}
\dfrac{\partial u}{\partial t} ( t, x ) = \alpha( t, x ) \dfrac{\partial^2 u}{\partial x^2}( t, x ) & 
\forall ( t, x ) \in [0,T] \times [ y, z ] \\
\text{IC:} & \\
u( 0, x ) = u_0( x ) & \forall x \in [ y, z ] \\
\text{BC:} & \\
u( t, y ) = v_1( t ) & \forall t \in [ 0, T ] \\
u( t, z ) = v_2( t ) & \forall t \in [ 0, T ]
\end{array}
\right.
\end{equation}

2. Related coefficients
```{r,eval=TRUE,message=FALSE, warning=FALSE}
# Diffusion parameter constant
Nt <- 300
Nx <- 150
```

3. Time grid
\begin{equation}
t_n = t_{n-1} + n h,
\qquad h = \frac{T - 0}{N_t}
\end{equation}

```{r,eval=TRUE,message=FALSE, warning=FALSE}
t0 <- 0
t1 <- 1
t <- cf_uniform_grid( t0, t1, Nt )
```

4. Space grid
\begin{equation}
x_0 = y, \qquad 
x_i = x_{i-1} + i g,\qquad 
g = \frac{z - y}{N_x}
\end{equation}

```{r,eval=TRUE,message=FALSE, warning=FALSE}
y <- -0.5
z <- 0.5
x <- cf_uniform_grid( y, z, Nx )
```

5. Discretization of parameters 
\begin{equation}
\alpha_{n,i} = \alpha( t_n, x_i )
\end{equation}

```{r,eval=TRUE,message=FALSE, warning=FALSE}
alpha <- matrix( 10^(-2.3), Nt, Nx )
```

5. Initial conditions
```{r,eval=TRUE,message=FALSE, warning=FALSE}
uf <- function( x ) if ( abs(x) <= 0.1 ) return( 1 ) else return( 0 )
uf <- Vectorize( uf )
u0 <- sapply( x, FUN = uf )
```

6. Boundary conditions
```{r,eval=TRUE,message=FALSE, warning=FALSE}
v1 <- rep( 0, Nt )
v2 <- rep( 0, Nt )
```
\begin{equation}
A = \begin{bmatrix}
b_1 & a_1 & 0 & 0 & \cdots & 0 \\
c_1 & b_2 & a_2 & 0 & \cdots & 0 \\
0 & c_2 & b_3 & a_3 & \cdots & 0 \\
\vdots & \vdots & \ddots & \ddots & \ddots & \vdots \\
0 & 0 & \cdots & c_{n-2} & b_{n-1} & a_{n-1} \\
0 & 0 & \cdots & 0 & c_{n-1} & b_n
\end{bmatrix}
\end{equation}

5. Euler implicit scheme has the following form for the equation
\begin{equation}
u_{n+1,i} - u_{n,i} = \alpha_{n,i} \frac{\Delta t_n}{\Delta x_{i}\ \Delta x_{i+1}}
( u_{n+1,i+1} - 2 u_{n+1,i} + u_{n+1,i-1} )
\end{equation}

From previous scheme at every time step $n$ we formulate a tridiagonal problem $A u_n = d$, with 
the folowing definitions.
\begin{eqnarray}
\lambda_{n, i} & = & \alpha_{n, i} \frac{\Delta t_n}{\Delta x_{i}\ \Delta x_{i+1}} \\
a_i & = & -\lambda_{n,i} \\
b_i & = & 1 + 2 \lambda_{n,i} \\
c_i & = & -\lambda_{n,i} \\
d_i & = & u_{n,i}
\end{eqnarray}

```{r,eval=TRUE,message=FALSE, warning=FALSE}
Ueu <- cf_diff_solv_euls( alpha, u0, v1, v2, t, x )
```
plot
```{r,eval=TRUE,fig.show='hold', fig.show='hold', fig.align="center", fig.width=7, fig.height=4}
plot_ly( x = Ueu$t, y = Ueu$x, z = Ueu$u, alpha = 0.8 ) %>% 
  layout( scene = list( xaxis = list(title = "t"),
                        yaxis = list(title = "x"),
                        zaxis = list(title = "u") ) ) %>%
  add_surface()
```

6. Solving with Crank-Nicolson implicit method
\begin{equation}
u_{n+1,i} + \theta (  \lambda^2_{n,i} \Delta_x u_{n+1,i+1} -  \lambda^1_{n,i} \Delta_x u_{n+1,i} ) =
u_{n,i} + ( 1 - \theta ) (  \lambda^2_{n,i} \Delta_x u_{n,i+1} -  \lambda^1_{n,i} \Delta_x u_{n,i} )
\end{equation}

\begin{eqnarray}
\lambda^1_{n,i} & = & \alpha_{n, i} \frac{\Delta t_n}{\Delta x_{i}\ \Delta x_{i+1}} \\
\lambda^2_{n,i} & = & \alpha_{n, i} \frac{\Delta t_n}{\Delta x_{i+1}\ \Delta x_{i+1}} \\
a_i & = & -\theta \lambda^1_{n,i} \\
b_i & = & 1 + \theta ( \lambda^1_{n,i} + \lambda^2_{n,i} ) \\
c_i & = & -\theta \lambda^2_{n,i} \\
d_i & = & u_{n,i} + ( 1 - \theta ) (  \lambda^2_{n,i} \Delta_x u_{n,i+1} -  \lambda^1_{n,i} \Delta_x u_{n,i} )
\end{eqnarray}

```{r,eval=TRUE,message=FALSE, warning=FALSE}
theta <- 0.25
Ucn <- cf_diff_solv_cns( theta, alpha, u0, v1, v2, t, x )
```
plot
```{r,eval=TRUE,fig.show='hold', fig.show='hold', fig.align="center", fig.width=7, fig.height=4}
plot_ly( x = Ucn$t, y = Ucn$x, z = Ucn$u, alpha = 0.8 ) %>% 
  layout( scene = list( xaxis = list(title = "t"),
                        yaxis = list(title = "x"),
                        zaxis = list(title = "u") ) ) %>%
  add_surface()
```

7. Plotting results
```{r,eval=TRUE,message=FALSE, warning=FALSE}
alph <- max( alpha )
phi <- function( x ) dnorm( x, mean = 0, sd = sqrt( 2 * alph * t[ Nt ] ) )
s <- sapply( x, FUN = function( y ) integrate( function( x, y ) uf( y - x ) * phi( x ), -Inf, Inf, y )$value )

eeu <- matrix( Ueu$u[Nt,] - s, Nx, 1 ) # Euler error
ecn <- matrix( Ucn$u[Nt,] - s, Nx, 1 )

```
